---
title: "Revisitation analysis"
# author: 
#   - name: "Scott Forrest"
#     url: https://swforrest.github.io/
#     orcid: 0000-0001-9529-0108
#     affiliation: Queensland University of Technology, CSIRO
#     email: "scottwforrest@gmail.com"
date: today
execute: 
  cache: false
bibliography: ../paperpile_references.bib
toc: true
number-sections: false
format: 
  html:
    self-contained: true
    code-fold: show
    code-tools: true
    df-print: paged
    code-line-numbers: true
    code-overflow: scroll
    fig-format: png
    fig-dpi: 300
  pdf:
    geometry: 
      - top=30mm
      - left=30mm
editor:
  source
abstract: |
  
---

## recurse package

The `recurse` R package can be used to analyse revisitation patterns in animal movement data. It provides functions to identify and quantify the number of times an animal returns to specific locations, as well as the duration of stays at those locations. This can be useful for understanding habitat use, foraging behavior, and territoriality.

#### Recurse vignette (tutorial)
<https://cran.r-project.org/web/packages/recurse/vignettes/recurse.html>

#### Key reference
Bracis, C., Bildstein, K. L., & Mueller, T. (2018). **Revisitation analysis uncovers spatio-temporal patterns in animal movement data**. Ecography, 41(11), 1801–1811. https://doi.org/10.1111/ecog.03618

#### Application paper

Campbell, H. A., Loewensteiner, D. A., Murphy, B. P., Pittard, S., & McMahon, C. R. (2020). **Seasonal movements and site utilisation by Asian water buffalo (Bubalus bubalis) in tropical savannas and floodplains of northern Australia**. Wildlife Research , 48(3), 230–239. https://doi.org/10.1071/WR20070


# Getting started

## Load packages

```{r}

# install.packages("recurse")

library(tidyverse)
library(sf)
library(move)
library(recurse)

```

## Load data

```{r}

### Load raw data
dingo_data <- read_csv("../data/tanami_collars.csv") %>% filter(x > 0)
dingo_data_sf <- st_as_sf(dingo_data, coords = c("x", "y"), crs = 4326)

# reproject the data into a projected CRS (UTM zone 52)
dingo_data_sf <- st_transform(dingo_data_sf, crs = 28352)

# convert back to dataframe
dingo_data <- dingo_data_sf %>%
  mutate(x = st_coordinates(geometry)[,1],
         y = st_coordinates(geometry)[,2]) %>%
  st_set_geometry(NULL)

# create a timestamp column
dingo_data$timestamp <- as.POSIXct(paste(dingo_data$gmt_date, dingo_data$gmt_time), 
                                   format = "%d/%m/%Y %H:%M:%S",
                                   tz = "UTC")

# convert to local time
dingo_data$timestamp <- with_tz(dingo_data$timestamp, tzone = "Australia/Darwin")

```

## Recurse analysis

```{r}

# select columns for recurse
dingo_recurse <- dingo_data %>% 
  mutate(id = dogname,
         t = timestamp) %>%
  dplyr::select("x", "y", "timestamp", "id") 

# convert to dataframe
dingo_recurse <- as.data.frame(dingo_recurse)

# calculate revisitations 
dingo_recurse_object <- getRecursions(dingo_recurse, radius = 500, 
                                     timeunits = "hours") 

# extract the revisitation stats
revisit_stats <- dingo_recurse_object$revisitStats

```

### Plot revisitation stats

```{r}

# plot the revisitations
plot(dingo_recurse_object, dingo_recurse)

hist(revisit_stats$timeInside[revisit_stats$timeInside < 100], 
     breaks = 100,
     main = "Time inside (hours)")

hist(revisit_stats$timeSinceLastVisit[revisit_stats$timeSinceLastVisit < 1000], 
     breaks = 100,
     main = "Time since last visit (hours)")

```


